\documentclass[11pt,a4paper]{article}

\usepackage[T1]{fontenc}
\usepackage{mathpazo}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{url}
\usepackage{booktabs}

\definecolor{linkblue}{RGB}{0,0,140}
\hypersetup{colorlinks=true, linkcolor=linkblue, urlcolor=linkblue, citecolor=linkblue}

\titleformat{\section}{\large\bfseries}{\thesection.}{0.5em}{}
\titlespacing{\section}{0pt}{1.5em}{0.8em}
\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection}{0.5em}{}

\setlength{\parindent}{0pt}
\setlength{\parskip}{0.6em}

\pagestyle{fancy}
\fancyhf{}
\rfoot{\thepage}
\renewcommand{\headrulewidth}{0pt}

\begin{document}

\begin{center}
{\LARGE \textbf{Samhati: A Peer-to-Peer Open Intelligence Network}}

\vspace{1em}

Mrunal Pendem\\
\texttt{samhati@proton.me}\\
\url{https://github.com/mrunalpendem123/Samhati}

\vspace{0.5em}
\textit{March 2026}
\end{center}

\vspace{1em}

% ══════════════════════════════════════════
\begin{abstract}
Only 1.2 billion of 8.1 billion people have used AI tools. The United States has 20,000 times more AI servers per capita than a low-income country. Three companies---OpenAI, AWS, and Microsoft---hold 84\% of the AI market. ChatGPT is banned in approximately 20 countries. The best models are closed. The infrastructure costs \$600 billion per year. The intelligence is metered, censored, and concentrated.

Meanwhile, open-weight small language models with 3--14 billion parameters now match frontier models on domain-specific tasks. A swarm of diverse small models outperforms GPT-4o on AlpacaEval. These models run on consumer hardware. What is missing is a way to connect them.

We propose Samhati, a peer-to-peer network where every device that uses AI also provides AI. Queries fan out to multiple nodes running specialized small models. Nodes peer-rank each other's answers via BradleyTerry aggregation. Quality is enforced by competition, not capital. No staking, no special hardware, no permission to join. Open intelligence for everyone, forever.
\end{abstract}

\vspace{1em}

% ══════════════════════════════════════════
\section{Introduction}

\subsection{The Cost of Thinking}

AI inference is a metered utility. OpenAI generated \$20 billion in annual recurring revenue in 2025, tripling from the prior year~\cite{openai-revenue}. The five largest hyperscalers---Amazon, Microsoft, Google, Meta, and Oracle---will spend over \$600 billion on AI infrastructure in 2026, a 36\% increase over 2025, with approximately 75\% directed at GPU servers and data centers~\cite{hyperscaler-capex}. NVIDIA captures roughly 90\% of AI accelerator spending~\cite{hyperscaler-capex}.

This infrastructure serves a pay-per-think model. API costs for frontier models range from \$3 to \$168 per million output tokens~\cite{ai-pricing}. Consumer subscriptions cost \$20--200 per month per service~\cite{openai-pricing}. Power users managing multiple AI tools spend \$100--200 monthly~\cite{subscription-fatigue}. The cost of intelligence scales linearly with usage.

The trajectory is clear: as AI becomes essential for work, education, medicine, and code, the people who cannot afford access are excluded from the most transformative technology of our generation.

\subsection{The Access Problem}

The exclusion is not hypothetical. Of 8.1 billion people on earth, only 1.2 billion have used AI tools---nearly 4 billion lack the basic prerequisites~\cite{world-bank}. High-income countries account for 87\% of notable AI models, 86\% of AI startups, and 91\% of venture capital funding, while representing just 17\% of the global population~\cite{csis-divide}. Africa holds less than 1\% of global data center capacity despite being home to 18\% of the world's population~\cite{csis-divide}. In some African markets, the cost of a single high-end GPU rivals the annual income of an average citizen~\cite{carnegie-africa}.

Access is not only limited by cost---it is limited by policy. ChatGPT is banned or unavailable in approximately 20 countries, including China, Russia, Iran, and several African nations~\cite{chatgpt-banned}. All three major Chinese chatbots strictly align with Beijing's official narratives on political matters, with censorship applied at both training and alignment stages~\cite{china-censorship}. Legal frameworks in at least 21 countries mandate or incentivize platforms to deploy machine learning to remove political, social, or religious speech~\cite{freedom-house}.

When three companies control 84\% of the AI market~\cite{ai-monopoly}, and 94\% of IT leaders report fear of vendor lock-in~\cite{vendor-lockin}, intelligence is not open---it is rented.

\subsection{Open Models Exist. Running Them Does Not Scale.}

The open-source community has produced remarkable models. Meta's Llama~3 405B approaches GPT-4 quality~\cite{llama3}. DeepSeek-R1 scored 89.1\% on MMLU, exceeding OpenAI's o1~\cite{deepseek-stats}. Qwen~2.5 achieved 79.7 on MMLU across model sizes from 0.5B to 72B~\cite{qwen25}.

But running these models locally is not feasible for individuals. Llama~3 405B at FP16 requires approximately 810GB of VRAM. Even a 70B model at Q4 quantization needs 35--40GB of RAM. This hardware costs tens of thousands of dollars and consumes kilowatts of power.

Open-weight models democratized the \textit{knowledge} of AI. What remains centralized is the \textit{compute} to run it.

\subsection{The Case for Small Language Models}

The answer is not bigger models on bigger hardware. It is smaller, specialized models working together.

Microsoft Phi-3-small (7B parameters) outperforms GPT-3.5-Turbo across language, reasoning, coding, and math benchmarks~\cite{phi3}. Mistral~7B outperformed Llama~2 13B on every benchmark at launch~\cite{mistral}. Liquid AI's foundation models, built on continuous-time dynamical systems rather than transformers, outperform models of similar and larger size~\cite{liquid}.

Most critically, ensembles of small models beat large ones. The Mixture-of-Agents approach~\cite{moa}, using only open-source models (Qwen, LLaMA, Mixtral), achieved 65.1\% on AlpacaEval 2.0---compared to GPT-4o's 57.5\%. Attention-MoA~\cite{attention-moa} extended this result, with an ensemble of small open-source models outperforming Claude-4.5-Sonnet and GPT-4.1 on MT-Bench.

The implication is profound: a network of diverse 3--14B parameter models, each specialized in its domain, can collectively produce intelligence that exceeds any single frontier model. This is the foundation of Samhati.

% ══════════════════════════════════════════
\section{Swarm Inference}

A query enters the network and is classified by complexity: Easy, Medium, or Hard. Easy and Medium queries dispatch to 3 nodes. Hard queries dispatch to 5 nodes and include a debate round~\cite{debate}.

Each node runs inference independently using \texttt{llama.cpp} and returns an answer alongside a Proof of Inference. Fan-out is parallel---all nodes begin simultaneously. Model heterogeneity is encouraged: a round might include Qwen-7B, Llama-8B, and DeepSeek-Coder-7B answering the same question.

For hard queries, a debate round follows initial inference: each node sees all other answers and rewrites its own. Multi-agent debate improves accuracy by +12.8 percentage points on mathematical reasoning~\cite{debate}.

% ══════════════════════════════════════════
\section{Peer Ranking}

Each pair of answers is evaluated by a randomly assigned judge node. The judge is instructed: \textit{``Judge only on factual correctness, logical reasoning, and completeness. Ignore confidence, style, and length.''} Judges cannot evaluate their own answers.

Assignment is randomized per round using $\text{BLAKE3}(\text{timestamp} \| \text{prompt} \| N)$, preventing collusion. For $N$ nodes, $\binom{N}{2}$ pairs are each judged by exactly one node. This is both more secure (colluders cannot control assignments) and more efficient (fewer ranking calls than exhaustive evaluation).

% ══════════════════════════════════════════
\section{BradleyTerry Aggregation}

Pairwise preferences are aggregated using the BradleyTerry model~\cite{bt} via Majorization-Minimization~\cite{mm}. Each node $i$ has latent strength $\beta_i$:

\begin{equation}
P(i \succ j) = \frac{e^{\beta_i}}{e^{\beta_i} + e^{\beta_j}}
\end{equation}

The algorithm iterates until $\max|\Delta\beta| < 10^{-6}$, typically in 10--25 iterations. Rankings from higher-ELO judges carry more weight. The winner is $\arg\max_i \beta_i$.

% ══════════════════════════════════════════
\section{ELO Reputation}

Each node maintains an ELO rating~\cite{elo}, starting at 1500:

\begin{equation}
\Delta = K \times \left(\text{actual} - \frac{1}{1 + 10^{(R_B - R_A)/400}}\right)
\end{equation}

$K=32$ for nodes with fewer than 1000 rounds; $K=16$ for established nodes. Floor is 100. Failed proof verification triggers a 200-point slash.

Domain-specific ELO is tracked separately: a node rated 1800 for Code and 1400 for Math will be selected for coding queries but not mathematical ones. ELO determines both node selection and ranking weight in BradleyTerry.

% ══════════════════════════════════════════
\section{Proof of Inference}

Each node produces a cryptographic proof from its output logprobs---the top-$K$ token probability distributions at each generation step. Tokens are grouped into 32-token chunks and BLAKE3-hashed from deterministic byte serialization (big-endian, sorted by token ID). The hash chain is Ed25519-signed with the node's private key.

The verifier checks: model hash, token count, chunk count, sequential indices, 5-minute freshness, registered \texttt{node\_pubkey}, and valid Ed25519 signature.

\subsection{Why Proof of Inference, Not Proof of Stake}

Existing decentralized AI networks require token staking~\cite{fortytwo}. Nodes must purchase tokens before participating. This creates a capital barrier that excludes the people an open intelligence network should serve: the student in India with a laptop but no cryptocurrency, the researcher in Nigeria with a GPU but no exchange access.

Staking optimizes for capital, not quality. A well-funded node with a bad model earns the right to participate by spending money.

Proof of Inference optimizes for compute: you must actually run a model to produce valid logprobs. The barrier is a laptop, not a wallet. Quality is enforced by peer ranking (Section~3), not by economic deposit.

\subsection{What Proof of Inference Is Not}

It is not TOPLOC~\cite{toploc}---it does not hash internal layer activations or verify which specific model ran. It verifies that \textit{a} model produced specific output distributions, signed by a specific node, within a recent time window. This is a weaker guarantee than activation-level verification but a stronger guarantee than staking, and it runs on any hardware without modification to the inference engine.

% ══════════════════════════════════════════
\section{Network}

Nodes discover each other through Solana~\cite{solana} on-chain registration and iroh~\cite{iroh} QUIC gossip.

A single Ed25519 keypair derives the node's Solana address (base58), iroh NodeId (hex), and proof signing key. On startup, the node registers on Solana, reads all existing NodeAccount PDAs, connects via iroh QUIC with relay-assisted NAT traversal, and broadcasts its model and endpoint every 10 seconds. Peers are verified against Solana registration before acceptance.

% ══════════════════════════════════════════
\section{Incentive}

Nodes earn SMTI tokens per round. The winner receives 60\% of emission; the remaining 40\% is distributed by ELO weight. Domain specialists (Code, Math, Reasoning) earn $1.5\times$ on matching queries.

On-chain demand counters track query domains in real time: if 42\% of queries are Code, running a Code specialist earns more. The network self-balances through economic signal.

There is no staking. Any device joins and earns immediately. Economic defense against Sybil attacks comes from compute cost: valid proofs require real inference.

% ══════════════════════════════════════════
\section{Settlement}

Each round settles on Solana via the \texttt{submit\_round} instruction. The on-chain record includes participant keys, BLAKE3 proof hashes, signed ELO deltas, winner, domain, and emission.

\begin{center}
\begin{tabular}{@{}lll@{}}
\toprule
\texttt{ProtocolConfig} & \texttt{[b"config"]} & authority, emission, demand \\
\texttt{NodeAccount} & \texttt{[b"node", pubkey]} & ELO, model, rounds, rewards \\
\texttt{RoundAccount} & \texttt{[b"round", id]} & participants, proofs, deltas \\
\bottomrule
\end{tabular}
\end{center}

% ══════════════════════════════════════════
\section{Security}

\textbf{Sybil.} Fake nodes must run real inference (Proof of Inference requires valid logprobs). Low-quality answers lose ELO and stop being selected. Compute cost exceeds reward.

\textbf{Collusion.} Judge assignment is $\text{BLAKE3}(\text{timestamp} \| \text{prompt} \| N)$, unpredictable and per-round. Colluders cannot control which pairs they judge.

\textbf{Replay.} Proofs expire in 5 minutes. Each is bound to a unique \texttt{node\_pubkey} via Ed25519.

\textbf{Freeloading.} A 0.5B model consistently loses to 7B models in ranking. ELO decays. Selection drops. The system self-corrects.

% ══════════════════════════════════════════
\section{Conclusion}

Three companies control 84\% of AI. Only 1.2 billion of 8.1 billion people have used it. ChatGPT is banned in 20 countries. The infrastructure costs \$600 billion per year. Meanwhile, a swarm of open-source small models outperforms GPT-4o.

Samhati connects these models into a collective intelligence. Every device that uses AI also provides it. Quality is enforced by competition---the best answer wins, regardless of the operator's capital. No staking, no special hardware, no permission required.

Open intelligence for everyone, forever.

% ══════════════════════════════════════════
\begin{thebibliography}{25}

\bibitem{openai-revenue}
PYMNTS, ``OpenAI's Annual Recurring Revenue Tripled to \$20 Billion in 2025,'' February 2026.

\bibitem{hyperscaler-capex}
IEEE ComSoc, ``Hyperscaler CapEx to Exceed \$600B in 2026, a 36\% Increase Over 2025,'' December 2025.

\bibitem{ai-pricing}
AI Empire Media, ``AI Pricing 2026: Costs Drop 40--70\%,'' 2026.

\bibitem{openai-pricing}
OpenAI, ``Pricing,'' \url{https://openai.com/pricing}, 2025.

\bibitem{subscription-fatigue}
AiZolo, ``All-in-One AI Subscription: Cut Costs by \$1,000+,'' 2025.

\bibitem{world-bank}
World Bank, ``Digital Progress and Trends Report 2025.''

\bibitem{csis-divide}
CSIS, ``From Divide to Delivery: How AI Can Serve the Global South,'' 2025.

\bibitem{carnegie-africa}
Carnegie Endowment, ``DeepSeek R1: Implications for Africa,'' March 2025.

\bibitem{chatgpt-banned}
Visual Capitalist, ``Mapped: Where ChatGPT is Banned in 2025,'' 2025.

\bibitem{china-censorship}
Carnegie Endowment, ``China's AI-Empowered Censorship: Strengths and Limitations,'' March 2026.

\bibitem{freedom-house}
Freedom House, ``The Repressive Power of Artificial Intelligence,'' Freedom on the Net 2023.

\bibitem{ai-monopoly}
TRENDS Research, ``How Big Tech's Monopoly of AI Threatens Fair Competition,'' 2025.

\bibitem{vendor-lockin}
GlobeNewsWire/Parallels, ``94\% of IT Leaders Fear Vendor Lock-In,'' February 2026.

\bibitem{llama3}
Meta AI, ``Llama 3,'' \url{https://ai.meta.com/blog/meta-llama-3/}, April 2024.

\bibitem{deepseek-stats}
SQ Magazine, ``DeepSeek AI Statistics 2026,'' 2026.

\bibitem{qwen25}
Qwen Team, ``Qwen2.5 Technical Report,'' \textit{arXiv:2412.15115}, December 2024.

\bibitem{phi3}
Microsoft Research, ``Phi-3 Technical Report,'' \textit{arXiv:2404.14219}, April 2024.

\bibitem{mistral}
A.~Jiang et al., ``Mistral 7B,'' \textit{arXiv:2310.06825}, October 2023.

\bibitem{liquid}
Liquid AI, ``The Future of AI Isn't Just Bigger---It's Specialized and Hybrid,'' \url{https://www.liquid.ai/blog/the-future-of-ai-isnt-just-bigger-its-specialized-and-hybrid}, 2025.

\bibitem{moa}
Together AI, ``Mixture-of-Agents Enhances Large Language Model Capabilities,'' \textit{arXiv:2406.04692}, 2024.

\bibitem{attention-moa}
``Attention-MoA: Optimizing Mixture-of-Agents,'' \textit{arXiv:2601.16596}, January 2026.

\bibitem{debate}
Y.~Du et al., ``Improving Factuality and Reasoning in Language Models Through Multiagent Debate,'' \textit{arXiv:2305.14325}, 2023.

\bibitem{bt}
R.~A. Bradley and M.~E. Terry, ``Rank Analysis of Incomplete Block Designs,'' \textit{Biometrika}, 39(3/4), 324--345, 1952.

\bibitem{mm}
K.~Lange et al., ``Optimization Transfer Using Surrogate Objective Functions,'' \textit{J.~Comp.~Graph.~Stat.}, 9(1), 1--20, 2000.

\bibitem{elo}
A.~E. Elo, \textit{The Rating of Chessplayers, Past and Present}, Arco, 1978.

\bibitem{toploc}
PrimeIntellect, ``TOPLOC: Token-Level Proof of Compute,'' \textit{arXiv:2501.16007}, 2025.

\bibitem{fortytwo}
Fortytwo Network, ``Swarm Inference with Peer-Ranked Consensus,'' \textit{arXiv:2510.24801}, 2025.

\bibitem{iroh}
n0 Computer, ``iroh,'' \url{https://iroh.computer}, 2024.

\bibitem{solana}
Solana Foundation, ``Solana Whitepaper,'' \url{https://solana.com/solana-whitepaper.pdf}, 2020.

\end{thebibliography}

\end{document}
